# Layers

```{r , eval=T, echo=F}
klippy::klippy()
```

```{r, eval=T, echo=F}
teachDS::img_centering("https://www.economist.com/img/b/1280/755/90/media-assets/image/20211002_WOC517.png", width="90%")
```


  * Geometries: `geom_line`
  
    * Aesthetics: x, y, color; size

    
```{r}
source("support/time-trend.R")
```


## Time series data

  * package: lubridate
  
Basic data types do not have `time` or `date` type. Time/Date are class that is constructed on top of character data, like 
```{r}
"2020-02-01"
"2020-02-01 13:58:55"
```

  * All date/time data sources are normally input as character. It requires parsing for computer to under it is date/time.
  
```{r}
lubridate::ymd("2020-02-01")
lubridate::ymd_hms("2020-02-01 13:58:55") # UTC time zone by default
lubridate::ymd_hms("2020-02-01 13:58:55", tz="Asia/Taipei")
```


```{r}
date = lubridate::ymd(c("2020-02-01", "2020-04-01",
  "2020-06-01", "2020-09-01"))
data = list()
data$Britain <- 
  data.frame(
    date = date,
    y = c(0, 50, 80, 100)
  )
data$Spain <-
  data.frame(
    date = date,
    y = c(0, 32, 53, 103)
  )
data$Italy <-
  data.frame(
    date = date,
    y = c(0, 50, 60, 99)
  )
```

```{r}
ggplot(data=dataAll) +
  geom_line(
    aes(
      x=date,
      y=y
    )
  )
```


```{r}
ggplot(data=dataAll) +
  geom_line(
    aes(
      x=date,
      y=y,
      group=country
    )
  )
```


## Aesthetics inheritance

  * In the past, we have done **data inheritance** (i.e. data in `ggplot()`) and both **data/aesthetics inheritance** (i.e, data and mapping in `ggplot()`). Actually you can do only **aesthetics inheritance** as well.

```{r}
sizeInput = 2 #input$sizeInput
plot <- list()
plot$p1 <- {
  ggplot(
    mapping=aes(
      x=date,
      y=y
    ))+
  geom_line(
    data=data$Britain,
    color="#096fa1",
    size = sizeInput
  )+
  geom_line(
    data=data$Spain,
    color="#ad8c97",
    size = sizeInput
  )+
  geom_line(
    data=data$Italy,
    color = "#983d4d",
    size = sizeInput
  )
}
plot$p1
```

## Sequence of layers

### Line stroke

```{r}
ggplot(
    mapping=aes(
      x=date,
      y=y
    ))+
  geom_line(
    data=data$Britain,
    color="#096fa1",
    size = sizeInput
  ) + 
  geom_line( # the last geom will be on the top
    data=data$Britain,
    color="black",
    size = sizeInput
  )  
```

```{r}
prop = 0.6

ggplot(
    mapping=aes(
      x=date,
      y=y
    ))+
  geom_line( # the last geom will be on the top
    data=data$Britain,
    color="black",
    size = sizeInput
  ) +  
  geom_line(
    data=data$Britain,
    color="#096fa1",
    size = sizeInput*prop
  ) 

```

## Create your first geom

```{r}
geom_lineWithStroke_error <- function(data, sizeInput, prop) {
    geom_line( # the last geom will be on the top
      data=data,
      color="black",
      size = sizeInput
    ) +  
    geom_line(
      data=data,
      color="#096fa1",
      size = sizeInput*prop
    ) 
  }
```

<div class="alert alert-info">
```{r}
class(plot$p1) # ggplot class
```

  * a list of 9 that complete the definitions of a plot

```{r}
gm <- geom_line( # the last geom will be on the top
    data=data$Britain,
    color="black",
    size = sizeInput
  ) 
class(gm) # Layer class
```

  * an environment that defines only a layer of a specific geometric structure and its aesthetics. 
  
  * part of a ggplot class object.
  
***

`+` operator must have the preceding object a ggplot object.

</div>

```{r}
geom_lineWithStroke_error(
  data=data$Britain, sizeInput=2, prop=.6
)
```

  * The error comes from the function body. It uses `+` on two layer class objects. There is no ggplot object presented.
  
  * `ggplot()+...` will always return a ggplot object. So all of below are ggplot objects:
  
    * `ggplot()`, `ggplot()+geom_point(...)`, `ggplot()+geom_point(...)+geom_line(...)`
    

### Layer adding

Instead of using `+` on each layers (i.e. adding geom one after the other), you can put all geom layers as a list, and use `+` to add the list all at once.

Other than the conventional adding:
```{r}
ggplot(
    mapping=aes(
      x=date,
      y=y
    ))+
  geom_line( 
    data=data$Britain,
    color="black",
    size = sizeInput
  ) +  
  geom_line( # the last geom will be on the top
    data=data$Britain,
    color="#096fa1",
    size = sizeInput*prop
  ) 
```

We can:
```{r}
ggplot(
  mapping=aes(
    x=date,
    y=y
  )) +
  list(
    geom_line( 
      data=data$Britain,
      color="black",
      size = sizeInput
    ),  
    geom_line( # the last geom will be on the top
      data=data$Britain,
      color="#096fa1",
      size = sizeInput*prop
    ) 
  )

```

### New geom function

<div class="alert alert-info">
A geom function that delivers a mixture of multiple geoms must return the mixture as a list, avoiding using `+` inside the function body.
</div>

```{r}
geom_lineWithStroke_prototype <- function(data, sizeInput, prop) {
  list(
    geom_line( 
      data=data,
      color="black",
      size = sizeInput
    ), 
    geom_line( # the last geom will be on the top
      data=data,
      color="#096fa1",
      size = sizeInput*prop
    ) 
  )
}

```

```{r}

{
  ggplot(
    mapping=aes(
      x=date,
      y=y
    ))+
  geom_lineWithStroke_prototype(
    data=data$Britain, sizeInput, prop
  )+
  geom_lineWithStroke_prototype(
    data=data$Spain, sizeInput, prop
  )+
  geom_lineWithStroke_prototype(
    data=data$Italy, sizeInput, prop
  )
}
```

### dot-dot-dot

  * <https://adv-r.hadley.nz/functions.html?q=dot-#fun-dot-dot-dot>

```{r}
geom_lineWithStroke_prototype2 <- function(data, sizeInput, prop, color, stroke, size) {
  list(
    geom_line(
      data=data,
      color=stroke,
      size = sizeInput
    ), 
    geom_line(  # the last geom will be on the top
      data=data,
      color=color,
      size = sizeInput*prop
    ) 
  )
}
```

```{r}
{
  ggplot(
    mapping=aes(
      x=date,
      y=y
    ))+
  geom_lineWithStroke_prototype2(
    data=data$Britain, sizeInput=sizeInput, prop=prop,
    color="#096fa1",
    stroke="white"
  )+
  geom_lineWithStroke_prototype2(
    data=data$Spain, sizeInput=sizeInput, prop=prop,
    color="#ad8c97",
    stroke="white"
  )+
  geom_lineWithStroke_prototype2(
    data=data$Italy, sizeInput=sizeInput, prop=prop,
    color="#983d4d",
    stroke="white"
  )
}

```

***

`geom_lineWithStroke` is an extension to an existing `geom_line` function. It would be better to make the input argument inline with `geom_line` so as to keep other geom_line options, such as linetype.

```{r}
help(geom_line)
```

```{r}
geom_lineWithStroke <- function(
  mapping = NULL,
  data = NULL,
  stat = "identity",
  position = "identity",
  na.rm = FALSE,
  orientation = NA,
  show.legend = NA,
  inherit.aes = TRUE,
  # set up default makes your function easy to use
  stroke = "white",
  prop = 0.9,
  size = 2, 
  ...){
  list(
    geom_line(
      data=data,
      color=stroke,
      size = size
    ),
    geom_line( # the last geom will be on the top
      mapping = mapping,
      data = data,
      stat = stat,
      position = position,
      na.rm = na.rm,
      orientation = orientation,
      show.legend = show.legend,
      inherit.aes = inherit.aes,
      size = size*prop,
      ...)
  )
}
```

<div class="alert alert-info">
`...` is a special argument for function. It means "whatever input whose input name is not specified in function arguments.

  * it can be passed directly into other functions who use `...`.
  
  * it can be accessed via `list()`.

```{r}
mySum <- function(...){
  browser()
  argList <- c(...)
  sum(argList)
}
mySum(2,3)
```

```{r}
greeting <- function(name, ...){
  browser()
  argList <- list(...)
  extraGreeting <- ""
  if(length(argList)!=0){
    extraGreeting <- paste(" Your", names(argList), " is ", argList)
  }
  
  cat("Hi ", name, ". ", extraGreeting)
  
  return(xx)
}
greeting("John", age=33)
```

</div>

```{r}

{
  ggplot(
    mapping=aes(
      x=date,
      y=y
    ))+
  geom_lineWithStroke(
    data=data$Britain,
    color="#096fa1"
  )+
  geom_lineWithStroke(
    data=data$Spain,
    color="#ad8c97"
  )+
  geom_lineWithStroke(
    data=data$Italy,
    color="#983d4d"
  )
}

```

```{r}
{
  ggplot(
    mapping=aes(
      x=date,
      y=y
    ))+
  geom_lineWithStroke(
    data=data$Britain,
    color="#096fa1",
    linetype=2
  )+
  geom_lineWithStroke(
    data=data$Spain,
    color="#ad8c97", 
    linetype=3
  )+
  geom_lineWithStroke(
    data=data$Italy,
    color="#983d4d",
    stroke="black",
    size=5
  )
}

```

## Summary

  * Layers (of geoms) can be added on top of each other through `+` operator by either:
  
    * **ggplot object** `+ geom1 + geom2`; or
    
    * **ggplote object** `+ list(geom1, geom2)`
    
  * When create a new geom function that consists of multiple geoms, use `list(geom1, geom2)` as function return.
  
  * When writing a function input argument names:
  
    * if in definition specifically **WITHOUT** default: it means something necessary. Users can not ignore.
    
    * if in definition specifically **WITH** default: it means something necessary, but there is a good choice of default that users will be happy with it most of the time.
    
    * `...`: a flexible argument especially when our function body has a call to some other function that has `...`
    
## Exercise

```{r, eval=T, echo=F}
teachDS::img_centering("https://www.economist.com/img/b/600/653/90/sites/default/files/images/2021/09/articles/body/20210925_fnc327.png", width="90%")
```

```{r, eval=T, echo=F}
teachDS::img_centering("https://www.economist.com/img/b/1280/755/90/sites/default/files/images/2021/03/articles/main/20210403_woc329.png", width="90%")
```

```{r, eval=T, echo=F}
teachDS::img_centering("https://clauswilke.com/dataviz/visualizing_associations_files/figure-html/CO2-slopegraph-three-year-1.png", width="90%")
```

  * <https://clauswilke.com/dataviz/visualizing-associations.html#associations-paired-data>

